\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-11 notes/2loopNLSM.tex
some changes
cut graphs

\section{Introduction}
\section{Review}
 To set the stage, it is worth distinguishing between the algebraic properties of functionally symmetric kinematics that satisfy color-kinematics globally, and some of the other constructions that appear in the literature.

@@

Many of the 1-loop results presented in \sect{} are informed by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ in YM, NLSM, or bi-adjoint scalar theory can be decomposed in terms of the trace basis
\be
\mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
\ee
where $T^a$ are generators of the gauge group and, because of the cyclicity of the trace, one of the legs is held fixed so that the sum runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial amplitudes $A_n[...]$.
Any tree-level color factor can be converted into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder color factors
\be
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\ee
associated with the graph
\be
\HalfLadder
\ee
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\be
\mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
\ee
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$ terms, indicating that there are additional relations (the Kleiss-Kuijf relations) amongst the partial amplitudes \cite{Kleiss:1988ne}.\footnote{The fundamental BCJ relations further reduce the basis of partial amplitudes down to $(n-3)!$ \cite{BCJ}.}
Since the Jacobi identity can be used to express any tree-level color factor in terms of the half ladders, it is enough to specify the numerators of the half ladder diagrams in order to satisfy color-kinematics duality at tree level.
At 1-loop, every diagram can be related to one topology as well, in this case the $n$-gon, but the pattern does not persist to 2-loops.

@@

As part of attemtping to construct a color-kinematics dual
representation for $\mathcal{N}=0$ Yang--Mills at two loops, Bern,
Davies, and Nohle (BDN) \cite{Bern:2015ooa} lay out the following core
assumptions in searching for double-copyable integrand numerators (ordered according to the
reference):
\begin{enumerate}
\item Locality: Diagram ``numerators'' are polynomials in momenta and
  polarizations, \eg the diagrams only have poles given by the
  propagators of the diagrams.
\item Power-counting: The maximum power of the loop momenta in the
  numerator is bounded by Feynman-gauge Feynman rules.
\item Cubic diagrams: The only diagrams with nonzero numerators are
  those with strictly cubic vertices.  This is generally a requirement
  to properly define the kinematic Jacobi relations.
\item Relabeling : The diagram numerators are only a function of the
  diagram's topology and labeling.
\item Symmetries: The diagram numerators are invariant under the
  automorphisms of the diagrams \emph{including signs that compensate
    for color-factor sign changes}.
\item Unitarity cuts: The cuts of the ansatz must reproduce the
  physical unitarity cuts of the theory.
\item Color-kinematics duality: The numerators obey some form of the
  kinematic Jacobi relations.
\end{enumerate}
Point 1 is generally a matter of convenience: there are many examples
in which ``numerators'' with controlled rational contributions have
been found
\cite{Square,WeinzierlBCJLagrangian,Mogull:2015adi,FivePointN4BCJ,Johansson:2017bfl}.
However, without a guiding principle about what types of rational
terms to use, the size of ans\"{a}tze quickly get out of hand.  Point
2 was motivated at the time by the hope that color-dual numerators
could arise naturally out of a special gauge choice. \jm{I'm not sure
  that we should bring up point 2.  I don't follow it and I don't
  think Alex did either since this wouldn't be the most general
  ansatz.  Also, why doesn't point 5 cover point 4?}\ace{I paraphrased
  these out of \cite{Bern:2015ooa}, thus the inclusion of Point 2.
  For Points 4 and 5, definitely 4 $\not \Rightarrow$ 5.  In the other
  direction, I could imagine a situation in which we violate 4 (eg we
  need to specify a reference ordering) but that symmetries are still
  respected as long as you also perform the correct actions on the
  auxiliary data.} On the other hand, recent work by Ben-Shahar,
Garozzo, and Johansson \cite{Ben-Shahar:2022ixa} \ace{Are there other
  good examples?}  demonstrates that constructing Feynman rules which
produce manifestly color-dual numerators for the NMHV sector of
Yang--Mills requires introducing additional auxiliary fields,
including higher-spin modes, which could shift the expected power
counting in non-trivial ways.  Point 3 is a requirement of the BCJ
double-copy: currently there is no known way of incorporating
non-cubic diagrams into the kinematic Jacobi relations.  Points 4 and
5 will be elaborated on in \cref{sec:syms}, Point 6 in
\cref{sec:unitarity}, and Point 7 in \cref{sec:jac-rels}.

@@

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes in the product of trees, and
restricting the set of compatible diagrams to those whose color
factors reduce to the appropriate DDM element when summing over
\emph{only the uncut color contractions}.  In this situation, the
color becomes an overall factor on the entire equality and thus can be
ignored.  Cuts that are organized in this manner are known as
``color-ordered'' cuts, in the same sense as color-ordered tree
amplitudes.  Since the DDM basis (and dual Kleiss-Kuijf amplitude
basis \cite{Kleiss:1988ne}) have $(n-2)!$ elements, one generally
needs to evaluate that many separate color-orderings of the same cut.
As an example, consider the ``kissing triangle'' cut topology, one of
the factorization channels of the iterated unitarity cut, which has
two topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
  \KissingTriangles =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

Even beginning at one loop, we encounter classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
  \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\ni
\frac{\delta(\textcolor{cut-color}{k_4^2})}{\textcolor{ucp-color}{s\, k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
\end{equation}
in which one of the \textcolor{ucp-color}{uncut propagators} is set
equal to an \textcolor{cut-color}{external on-shell propagator}.
Discussions about how to resolve these types of singularities can be
found in Refs.~\cite{Edison:2022jln, Edison:2022smn, Edison:2020uzf,
  Geyer:2015bja, Bern:2012uf} \citepls{Minahan}.  In addition to the
bubble-on-external-leg cuts, there is a related class of cuts in which
an uncut propagator is set to zero by an \emph{internal cut
  condition}, for instance,
\begin{equation}
  \label{eq:other-path}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

@@

\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed} = (s_{13}-s_{12}) \epsilon_{(12)} \epsilon_{(34)}
\ee
One was to resolve this is to introduce higher spin modes in order to accomodate the internal index contractions needed to recover the s-channal contract:
\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{Bfield}{nhpRed}{}{nhpRed} = s_{12 }(\epsilon_{(13)} \epsilon_{(24)} - \epsilon_{(14)} \epsilon_{(23)})
\ee
While introducing succesively higher spiin states to the NMHV lagrangian constructed in \citepls{} could in principle work at tree level, we will see that it is not consiisntent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above. This is easily understood in terms of a polarization decomposition introduced by one of the authors \citepls{} that decomposes vector amplitudes in terms of building blocks that preserve partial amplitude relations. 
\subsection{Block Decomposition of Yang-Mills}
Describe how different orders in $\varepsilon_i\varepsilon_j$ in pure Yang-Mills captures different cubic and contact sectors of the theory. All the $D$-dimensional Lagnrangians above can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \citepls{}:
\begin{eBox}
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\end{eBox}
By a simple mass-dimensioin argument, we know the tree-level cubic sector is simply:
\be
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
and likewise at one-loop:
\be
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
We can select out both of these manifestly cubic sectors by selecting only $\epsilon^+\epsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$ theory is just the following blocks in the RABD expansion of Yang-Mills:
\be
A(...,\bar{J}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\ee
and similarly so at one-loop:
\be
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}
\ee
To recover NLSM at one-loop, we need to extract out the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars:
\begin{eBox}
\be
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\epsilon\rightarrow k}
\ee
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section. 
\subsection{Semi-abelian Yang-Mills theory}
\subsection{One-loop SDYM integrands}
\be
\simplebox = g_{\text{YM}}^4 X(k_1,l_1)X(k_2,l_2)X(k_3,l_3)X(k_4,l_4)
\ee
\subsection{One-loop NLSM Integrands}

Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.
\be
\simplebox = (2 f_\pi^{-1})^4 (l_1 k_1)(l_2 k_2)(l_3 k_3)(l_4 k_4)
\ee
where we define sequential loop momenta on the $n$-gon as $l_{i+1}= l_i+k_i$. Given this, the $n$-gon permits a equivalent expression in terms of inverse propagators:
\be
\simplebox =f_\pi^{-4} [12][23][34][41]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_i^L - l_R^2$
\subsection{Two-loop obstruction}

@@

Similar to the pion construction in \cref{sec:pions}, we use the
kinematic Jacobi relations to define the numerators for all cubic
graphs in terms of the two basis graphs, including allowing nontrivial
numerators on BEL diagrams.  We will simply avoid imposing constraints
that directly involve such pathological diagrams.  That leaves us with
14 non-pathological cubic diagrams, shown in \cref{fig:ym-mc}, on
which to impose maximal cuts and symmetries.  Doing so, we are left
with AA total parameters in the unrestricted ansatz and BB total in
the BDN-restricted ansatz.

\begin{figure}
  \foreach \x in {1,...,14}
  {
    \begin{minipage}{0.32\textwidth}
      \includegraphics[width=\textwidth]{temp-figs/nonpath-mc-\x.pdf}
    \end{minipage}
  }
  \caption{The 14 non-pathological maximal-cut diagrams}
  \label{fig:ym-mc}
\end{figure}

Proceeding to the next-to-maximal cuts, we continue to avoid
pathological diagrams including the newly-appearing type discussed in
\cref{eq:other-path}. After discarding these types of cuts, we are
left with 8 \ace{(maybe a few more to account for dangling trees)}
unique cut topologies, shown in \cref{fig:ym-nmc}, to check.  Seven of
these next-to-maximal cuts are consistent with the kinematic Jacobi
relations and symmetries.  Interestingly, the ``kissing triangles''
cut,
\begin{equation*}
  \KissingTriangles \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.

>>>>

\section{Introduction}
\section{Background}
 To set the stage, it is worth distinguishing between the algebraic properties of functionally symmetric kinematics that satisfy color-kinematics globally, and some of the other constructions that appear in the literature.

@@

Many of the 1-loop results presented in \sect{} are informed by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ in YM, NLSM, or bi-adjoint scalar theory can be decomposed in terms of the trace basis
\be
\mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
\ee
where $T^a$ are generators of the gauge group and, because of the cyclicity of the trace, one of the legs is held fixed so that the sum runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial amplitudes $A_n[...]$.
Any tree-level color factor can be converted into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder color factors
\be
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\ee
associated with the graph
\be
\halfLadder
\ee
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\be
\mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
\ee
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$ terms, indicating that there are additional relations (the Kleiss-Kuijf relations) amongst the partial amplitudes \cite{Kleiss:1988ne}.\footnote{The fundamental BCJ relations further reduce the basis of partial amplitudes down to $(n-3)!$ \cite{BCJ}.}
Since the Jacobi identity can be used to express any tree-level color factor in terms of the half ladders, it is enough to specify the numerators of the half ladder diagrams in order to satisfy color-kinematics duality at tree level.
At 1-loop, every diagram can be related to one topology as well, in this case the $n$-gon, but the pattern does not persist to 2-loops.

@@

As part of attemtping to construct a color-kinematics dual
representation for $\mathcal{N}=0$ Yang--Mills at two loops, Bern,
Davies, and Nohle (BDN) \cite{Bern:2015ooa} lay out the following core
assumptions in searching for double-copyable integrand numerators (ordered according to the
reference):
\begin{enumerate}
\item \textbf{Locality}: Diagram ``numerators'' are polynomials in momenta and
  polarizations, \eg the diagrams only have poles given by the
  propagators of the diagrams.
\item  \textbf{Power-counting:} The maximum power of the loop momenta in the
  numerator is bounded by Feynman-gauge Feynman rules.
\item  \textbf{Cubic diagrams}: The only diagrams with nonzero numerators are
  those with strictly cubic vertices.  This is generally a requirement
  to properly define the kinematic Jacobi relations.
\item  \textbf{Relabeling}: The diagram numerators are only a function of the
  diagram's topology and labeling.
\item  \textbf{Symmetries}: The diagram numerators are invariant under the
  automorphisms of the diagrams \emph{including signs that compensate
    for color-factor sign changes}.
\item  \textbf{Unitarity cuts}: The cuts of the ansatz must reproduce the
  physical unitarity cuts of the theory.
\item  \textbf{Color-kinematics duality}: The numerators obey some form of the
  kinematic Jacobi relations.
\end{enumerate}
Point 1 is generally a matter of convenience: there are many examples
in which ``numerators'' with controlled rational contributions have
been found
\cite{Square,WeinzierlBCJLagrangian,Mogull:2015adi,FivePointN4BCJ,Johansson:2017bfl}.
However, without a guiding principle about what types of rational
terms to use, the size of ans\"{a}tze quickly get out of hand.  Point
2 was motivated at the time by the hope that color-dual numerators
could arise naturally out of a special gauge choice. \jm{I'm not sure
  that we should bring up point 2.  I don't follow it and I don't
  think Alex did either since this wouldn't be the most general
  ansatz.  Also, why doesn't point 5 cover point 4?}\ace{I paraphrased
  these out of \cite{Bern:2015ooa}, thus the inclusion of Point 2.
  For Points 4 and 5, definitely 4 $\not \Rightarrow$ 5.  In the other
  direction, I could imagine a situation in which we violate 4 (eg we
  need to specify a reference ordering) but that symmetries are still
  respected as long as you also perform the correct actions on the
  auxiliary data.} On the other hand, recent work by Ben-Shahar,
Garozzo, and Johansson \cite{Ben-Shahar:2022ixa} \ace{Are there other
  good examples?}  demonstrates that constructing Feynman rules which
produce manifestly color-dual numerators for the NMHV sector of
Yang--Mills requires introducing additional auxiliary fields,
including higher-spin modes, which could shift the expected power
counting in non-trivial ways.  Point 3 is a requirement of the BCJ
double-copy: currently there is no known way of incorporating
non-cubic diagrams into the kinematic Jacobi relations.  Points 4 and
5 will be elaborated on in \cref{sec:syms}, Point 6 in
\cref{sec:unitarity}, and Point 7 in \cref{sec:jac-rels}.

@@

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes in the product of trees, and
restricting the set of compatible diagrams to those whose color
factors reduce to the appropriate DDM element when summing over
\emph{only the uncut color contractions}.  In this situation, the
color becomes an overall factor on the entire equality and thus can be
ignored.  Cuts that are organized in this manner are known as
``color-ordered'' cuts, in the same sense as color-ordered tree
amplitudes.  Since the DDM basis (and dual Kleiss-Kuijf amplitude
basis \cite{Kleiss:1988ne}) have $(n-2)!$ elements, one generally
needs to evaluate that many separate color-orderings of the same cut.
As an example, consider the ``kissing triangle'' cut topology, one of
the factorization channels of the iterated unitarity cut, which has
two topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

Even beginning at one loop, we encounter classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
\frac{\delta(\textcolor{cut-color}{k_4^2})}{\textcolor{ucp-color}{s\, k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
  \in   \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\end{equation}
in which one of the \textcolor{ucp-color}{uncut propagators} is set
equal to an \textcolor{cut-color}{external on-shell propagator}.
Discussions about how to resolve these types of singularities can be
found in Refs.~\cite{Edison:2022jln, Edison:2022smn, Edison:2020uzf,
  Geyer:2015bja, Bern:2012uf} \citepls{Minahan}.  In addition to the
bubble-on-external-leg cuts, there is a related class of cuts in which
an uncut propagator is set to zero by an \emph{internal cut
  condition}, for instance,
\begin{equation}
  \label{eq:other-path}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

@@

\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed} = (s_{13}-s_{12}) \epsilon_{(12)} \epsilon_{(34)}
One was to resolve this is to introduce higher spin modes in order to accomodate the internal index contractions needed to recover the s-channal contract:
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{Bfield}{nhpRed}{}{nhpRed} = s_{12 }(\epsilon_{(13)} \epsilon_{(24)} - \epsilon_{(14)} \epsilon_{(23)})
While introducing succesively higher spiin states to the NMHV lagrangian constructed in \citepls{} could in principle work at tree level, we will see that it is not consiisntent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above. This is easily understood in terms of a polarization decomposition introduced by one of the authors \citepls{} that decomposes vector amplitudes in terms of building blocks that preserve partial amplitude relations. 
\subsection{Block Decomposition of Yang-Mills}
Describe how different orders in $\varepsilon_i\varepsilon_j$ in pure Yang-Mills captures different cubic and contact sectors of the theory. All the $D$-dimensional Lagnrangians above can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \citepls{}:
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
By a simple mass-dimensioin argument, we know the tree-level cubic sector is simply:
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
and likewise at one-loop:
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
We can select out both of these manifestly cubic sectors by selecting only $\epsilon^+\epsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$ theory is just the following blocks in the RABD expansion of Yang-Mills:
A(...,\bar{J}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
and similarly so at one-loop:
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}
To recover NLSM at one-loop, we need to extract out the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars:
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\epsilon\rightarrow k}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section. 
\subsection{Semi-abelian Yang-Mills theory}
\subsection{One-loop color-dual integrands}
\simplebox = g_{\text{YM}}^4 X(k_1,l_1)X(k_2,l_2)X(k_3,l_3)X(k_4,l_4)
\nGonScalar
Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.
\simplebox = (2 f_\pi^{-1})^4 (l_1 k_1)(l_2 k_2)(l_3 k_3)(l_4 k_4)
where we define sequential loop momenta on the $n$-gon as $l_{i+1}= l_i+k_i$. Given this, the $n$-gon permits a equivalent expression in terms of inverse propagators:
\simplebox =f_\pi^{-4} [12][23][34][41]
Where we have define the antisymmetric kinematic variable, $[LR] = l_i^L - l_R^2$
\subsection{Two-loop obstruction}

@@

Similar to the pion construction in \cref{sec:pions}, we use the
kinematic Jacobi relations to define the numerators for all cubic
graphs in terms of the two basis graphs, including allowing nontrivial
numerators on BEL diagrams.  We will simply avoid imposing constraints
that directly involve such pathological diagrams.  That leaves us with
14 non-pathological cubic diagrams, shown in \cref{fig:ym-mc}, on
which to impose maximal cuts and symmetries.  Doing so, we are left
with AA total parameters in the unrestricted ansatz and BB total in
the BDN-restricted ansatz.
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH,
}
  \caption{The 14 non-pathological maximal-cut diagrams}
  \end{figure}

Proceeding to the next-to-maximal cuts, we continue to avoid
pathological diagrams including the newly-appearing type discussed in
\cref{eq:other-path}. After discarding these types of cuts, we are
left with 8 \ace{(maybe a few more to account for dangling trees)}
unique cut topologies, shown in \cref{fig:ym-nmc}, to check.  Seven of
these next-to-maximal cuts are consistent with the kinematic Jacobi
relations and symmetries.  Interestingly, the ``kissing triangles''
cut,
\begin{equation*}
   \NMCutC \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.
\begin{figure}
\centering
 { \NMCutH, \NMCutG,  \NMCutD,\NMCutB, 
 \\
 \NMCutE, \NMCutF, \NMCutA, \NMCutC,  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams.}
  \label{fig:ym-nmc}
\end{figure}
